Particle Interception as a Function of Reynolds Number and Collector Density

### 2020-08-29 code/analysis.Rmd
analysis.rmd is now on version 2

---
title: "Particle Interception as a Function of Reynolds Number and Collector Density"
author: "Jordan Wingenroth"
date: "8/26/2020"
toc: TRUE
output:
  pdf_document: default
---

@@

@@

length(pumpfiles)
length(trapfiles)

@@

As you see, there are different numbers of files for sediment trap versus peristaltic pump data.

@@

tibble(pumpdate, pumpdate %in% trapdate)

@@

tidypump <- lapply(seq_along(x), function(i) {
  select(x[[i]], 
         loc = Location, 
         ht = Height,
         t = `time series`,
         mvc = contains("(ppm)")) %>%
    filter(t < 21) %>% #filter a few timepoints outside the normal window 
    mutate(t = (t-min(t)+1)*300, #convert from timestep to seconds
           mvc = as.numeric(mvc), 
           date = as.numeric(names(x)[[i]])) %>%
    filter(!is.na(mvc))
  }
  )

@@

pump

@@

Let's plot the decay curves separately to see what the data looks like qualitatively.

pump %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_point(alpha = .5, size = .75) +
  facet_wrap(~date)

@@

length(unique(pump$date))

@@

@@

sum(unique(pump$date) %in% metadata$date)

Hoorah! We have metadata for all of our runs of interest.

@@

Now then, let's remove the biofouled runs, which won't be used in assessing our primary hypothesis about Reynolds number and collector density, and we'll see what we're left with.

pump %>%
  filter(growth_days==0) %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_point(alpha = .5, size = .75) +
  geom_rect(data = filter(pump, date == 190417), 
                          fill = NA, colour = "red", xmin = -Inf,xmax = Inf,
            ymin = -Inf,ymax = Inf) +
  facet_wrap(~date) +
  scale_x_continuous(breaks = c(0,3000,6000), limits = c(0,6000))

These runs do look quite familiar to me because I think this is about my 10th time through this analysis. `190417` is the run with clear issues, in the early timesteps more specifically. During one of those previous analyses, I found that there were several runs where **part** of the timespan was clearly out of line with the expected decay pattern. Fortunately we tracked metadata about the logistics of laboratory analysis, and it just so happens that the clearly erroneous data were all from samples processed by one researcher in particular. `190417` is the only run out of these 13 where they were involved in processing samples, at least since we started recording initials on our work. We repeated that run on `190506` and got results more in line with exponential decay.

So that leaves 13 runs. Aside from one run at 232 collectors/$\text m^2$, the result of a miscalculation during dowel installation, the 12 remaining are a perfect match for our parameter space of 4 densities (including 0 control) and 3 velocities/Re values.

@@

pumpfinal %>%
  ggplot(aes(x = t, y = mvc)) +
  geom_point(alpha = .5, size = .75) +
  facet_grid(pump_freq~dowel_density) +
  scale_x_continuous(breaks = c(0,3000,6000), limits = c(0,6000))

@@

trap
```

So, across 21 runs, we had 186 samples. We were set up to collect 9 per run, but traps broke, filters slipped and spilled sediment, et cetera. $186/21 =$ `r 186/21`, so our success rate was actually pretty good, although some of those measurements were probably erroneous despite being measured. Let's see what we're working with and then we can maybe try to remove outliers.

lm(sed~station, data = trap) %>% summary()

@@

So, downstream traps tended to have less settling. This challenges our assumption of uniformity throughout the test section and might justify the use of a spline, as Justin and Candace had suggested.

trap %>%
  group_by(date) %>%
  mutate(mean = mean(sed)) %>%
  arrange(mean) %>%
  ggplot(aes(x = factor(date, levels = unique(date)), y = sed)) +
  geom_boxplot() +
  labs(x = "date", y = "sediment mass (mg/trap)") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

On the whole, the sediment data seems pretty normal and independent. Let's zoom in to just the 12 runs used in the main analysis.

@@

trapfinal %>%
  group_by(date) %>%
  mutate(mean = mean(sed)) %>%
  arrange(mean) %>%
  ggplot(aes(x = factor(date, levels = unique(date)), y = sed)) +
  geom_boxplot() +
  labs(x = "date", y = "sediment mass (mg/trap)") +
  scale_x_discrete(guide = guide_axis(n.dodge = 2))

@@

A pretty broad range, without any clear outliers to knock out, at least by my sense of it. The whiskers are limited to $1.5 \cdot \text{IQR}$ each (both up and down), so while there are 4 points outside that range for their respective runs, they're definitely all within the range of the dataset as a whole.

@@

@@

So then, k_s as a function of our independent vars:

@@

fitdata[fitdata$date==190802,"t"] <- (fitdata[fitdata$date==190729,"t"] + 
                                        fitdata[fitdata$date==190321,"t"])/2

fitdata
```

@@

final[final$date==190802,"k_s"] <- (final[final$date==190729,"k_s"] + 
                                      final[final$date==190321,"k_s"])/2

@@

- i did find a way to propogate error through these calculations quite easily
    - I think including error bars on our main graph (eta over Rey) might be a good idea?
    - Is it irresponsible to do so with `n = 1` per treatment?

>>>>

---
title: "Particle Interception as a Function of Reynolds Number and Collector Density"
subtitle: "Version 2.0"
author: "Jordan Wingenroth"
date: "8/28/2020"
toc: TRUE
output:
  pdf_document: default
---

@@

### Changes from Version 1 (8/26)

I pared back this section to reduce the page count. Please refer to the original version to see visualizations and more detailed steps and notes.

### Data background

@@

@@

@@

@@

tidypump <- lapply(seq_along(x), function(i) {
  select(x[[i]], 
         loc = Location, 
         ht = Height,
         t = `time series`,
         mvc = contains("(ppm)")) %>%
    filter(t < 21) %>% #filter a few timepoints outside the normal window 
    mutate(t = (t-min(t)+1)*300, #convert from timestep to seconds
           mvc = as.numeric(mvc), 
           date = as.numeric(names(x)[[i]])) %>%
    filter(mvc<80, mvc>8) #outliers were removed based on the residual graph
  }
  )

@@

@@

@@

@@

```{r message=FALSE}

@@

@@

Now then, let's remove the biofouled runs and a couple runs with other issues, which won't be used in assessing our primary hypothesis about Reynolds number and collector density, and we'll see what we're left with.

@@

@@

@@

So, across 21 runs, we had 186 samples. We were set up to collect 9 per run, but traps broke, filters slipped and spilled sediment, et cetera. $186/21 =$ `r 186/21`, so our success rate was actually pretty good, although some of those measurements were probably erroneous despite being measured.

@@

@@

@@

### NEW ADDITION: Uncertainty propogation

Here is a nice function found on StackExchange.com, which performs error propogation in R using derivatives:

mutate_with_error = function(.data, f) {
  exprs = list(
    # expression to compute new variable values
    deparse(f[[3]]),

    # expression to compute new variable errors
    sapply(all.vars(f[[3]]), function(v) {
      dfdp = deparse(D(f[[3]], v))
      sprintf('(d%s*(%s))^2', v, dfdp)
    }) %>%
      paste(collapse='+') %>%
      sprintf('sqrt(%s)', .)
  )
  names(exprs) = c(
    deparse(f[[2]]),
    sprintf('d%s', deparse(f[[2]]))
  )

  .data %>%
    # the standard evaluation alternative of mutate()
    mutate_(.dots=exprs)
}

@@

So then, k_tot as a function of our independent vars:

@@

# THIS CODE WAS USED IN THE FIRST VERSION BUT SHOULD BE LEFT OUT OF ERROR PROPOGATION VERSION

# 
# fitdata[fitdata$date==190802,"t"] <- (fitdata[fitdata$date==190729,"t"] + 
#                                         fitdata[fitdata$date==190321,"t"])/2
# 
# fitdata

@@

# THIS CODE WAS USED IN THE FIRST VERSION BUT SHOULD BE LEFT OUT OF ERROR PROPOGATION VERSION

# final[final$date==190802,"k_s"] <- (final[final$date==190729,"k_s"] + 
#                                       final[final$date==190321,"k_s"])/2

@@
